The Nature and Orbit of the Ophiuchus Stream

### 2014-12-31 tex/ms.tex
Added Imants Platais and Rosemary Wyse as co-authors and implemented their comments. Implemented Christian's comments (provided by Melissa Ness). Expanded the Conclusions section a bit.

\author{Branimir Sesar\altaffilmark{1,2}}
\author{Jo Bovy\altaffilmark{3,4}}
\author{Edouard J.~Bernard\altaffilmark{5}}
\author{Nelson Caldwell\altaffilmark{6}}
\author{Judith G.~Cohen\altaffilmark{7}}
\author{Morgan Fouesneau\altaffilmark{1}}
\author{Christian I.~Johnson\altaffilmark{6}}
\author{Melissa Ness\altaffilmark{1}}
%\author{Maria Bergemann\altaffilmark{1}}
\author{Annette M.~N.~Ferguson\altaffilmark{5}}
\author{Nicolas F.~Martin\altaffilmark{8,1}}
\author{Hans-Walter Rix\altaffilmark{1}}
\author{Edward F.~Schlafly\altaffilmark{1}}
%\author{builders}
\author{W.~S.~Burgett\altaffilmark{15}}
\author{K.~C.~Chambers\altaffilmark{9}} 
%\author{L.~Denneau\altaffilmark{9}} 
%\author{P.~Draper\altaffilmark{10}} 
%\author{H.~Flewelling\altaffilmark{9}} 
%\author{T.~Grav\altaffilmark{11}} 
%\author{J.~N.~Heasley\altaffilmark{9}} 
\author{K.~W.~Hodapp\altaffilmark{9}} 
%\author{M.~E.~Huber\altaffilmark{9}} 
%\author{R.~Jedicke\altaffilmark{9}} 
\author{N.~Kaiser\altaffilmark{9}} 
%\author{R.-P.~Kudritzki\altaffilmark{9}}
%\author{G.~A.~Luppino\altaffilmark{9}}
%\author{R.~H.~Lupton\altaffilmark{12}}
\author{E.~A.~Magnier\altaffilmark{9}}
%\author{N.~Metcalfe\altaffilmark{10}}
%\author{D.~G.~Monet\altaffilmark{13}}
%\author{J.~S.~Morgan\altaffilmark{9}}
%\author{P.~M.~Onaka\altaffilmark{9}}
%\author{P.~A.~Price\altaffilmark{12}}
%\author{C.~W.~Stubbs\altaffilmark{14}}
%\author{W.~Sweeney\altaffilmark{9}}
\author{J.~L.~Tonry\altaffilmark{9}}
%\author{R.~J.~Wainscoat\altaffilmark{9}}
\author{C.~Waters\altaffilmark{9}}

@@

\begin{abstract}
The Ophiuchus stream is the most recently discovered stellar tidal stream in the
Milky Way \citep{ber14b}. We present high-quality spectroscopic data for 14
stream member stars obtained using Keck and MMT telescopes. We confirm the
stream as a fast moving ($v_{los}\sim290$ km s$^{-1}$), kinematically-cold group
($\sigma_{v_{los}}\lesssim1$ km s$^{-1}$) of $\alpha-$enhanced and metal-poor
stars (${\rm [\alpha/Fe]\sim0.4}$ dex, ${\rm [Fe/H]\sim-2.0}$ dex). Using a
probabilistic technique, we model the stream simultaneously in line-of-sight
velocity, color-magnitude, coordinate, and proper motion space. We find that
that the stream extends from 8 to 9.5 kpc from the Sun and that its deprojected
length is $\sim1.6$ kpc. The analysis of the stellar population contained in the
stream suggests that it is $\sim13$ Gyr old, and that its initial stellar mass
was $\sim2\times10^4$ $M_\sun$ (or at least $\ga4\times10^3$ $M_\sun$). We do
not detect a significant overdensity of stars along the stream that would
indicate the presence of a progenitor, and conclude that the stream is all that
is left of the progenitor. Assuming a potential for the Milky Way that is
consistent with a large variety of dynamical constraints, we fit the orbit of
the stream. We find that the stream has an orbital period of $\sim360$ Myr, and
is on a fairly eccentric orbit ($e\sim0.68$) with a pericenter of $\sim3.5$ kpc
and an apocenter of $\sim17.5$ kpc. Based on the available information, we
conclude that the progenitor of the stream was a globular cluster and estimate
the time of disruption at $\sim250$ Myr ago.
\end{abstract}

@@

To address the above questions, we need to know the orbit of the Ophiuchus
stream, and to determine its orbit we need to measure the stream's line-of-sight
velocity, distance, and proper motion. In Section~\ref{data}, we describe the
data we use in this work; the PS1 photometry and astrometry, the spectroscopic
follow-up of candidate stream members, and the measurement of their
line-of-sight velocities, chemical abundances, and proper motions. In
Section~\ref{results}, we provide a detailed characterization of the stream in
position, velocity, and abundance (7-D) phase space. The constraints obtained in
Section~\ref{results} are then used to constrain and examine the orbit of the
stream (Section~\ref{orbit}). In Section~\ref{conclusions}, we discuss the
nature of the stream's peculiar orbit, highlight the solved and uncovered
puzzles related to the stream, and finally present our conclusions.

@@

\subsection{Chemical abundances}

Even though the primary goal of spectroscopic observations was to obtain precise
radial velocities, the wavelength range and the resolution of Hectochelle
spectra are sufficient to allow estimates of chemical abundances.

We determined stellar parameters from the continuum-normalised, radial
velocity-corrected spectra using the SMH code of \citet{cas14}, which is built
on the MOOG code of \citet{sne73}. Kurucz model atmospheres \citep{ck04} and a
line list compiled from \citet{fre10} and \citet{yon05} by \citet{cas14} were
used. First, effective temperatures were calculated from dereddened PS1 $g_{P1}$
and $r_{P1}$ bands\footnote{Transformed to the Sloan Digital Sky Survey (SDSS;
\citealt{yor00}) $g$ and $r$ bands using relations of \citet{ton12}.} using
Equation 3 of \citet{ive08} and spectroscopic temperatures were optimized around
this value using the SMH code, by removing abundance trends with line excitation
potential. Parameters of $\log g$ and ${\rm [Fe/H]}$ were determined using 14 Fe
I and 2-3 Fe II lines to achieve ionization balance, and microturbulence was
calculated by removing abundance trends with reduced equivalent width of the
lines. Estimates of the $\alpha-$enhancement were obtained using only the few
available clean Mg, Ca and Ti (I and II) lines, which comprised a total of 6-8
lines per star.

@@

Proper motions are indispensable constraints for the orbit of a stream (e.g.,
\citealt{kop10}) To measure the proper motion of stars in the vicinity of the
Ophiuchus stream, we combine the astrometry provided by USNO-B \citep{mon03} and
2MASS \citep{skr06} catalogs with the PS1 catalog. The USNO-B catalog lists
photometry and astrometry measured from photographic plates in five different
band-passes ($O$, $E$, $J$, $F$, and $N$). The USNO-B plates were exposed at
different epochs, and thus each object in the catalog can have a maximum of five
recorded positions. The 2MASS catalog provides only one position entry per
object.

Since these catalogs are not tied to the same astrometric reference system, we
first calibrate USNO-B and 2MASS positions to the reference frame defined by
positions of galaxies observed in PS1. We define galaxies as objects that have
the difference between Point-Spread-Function (PSF) and aperture magnitudes in
PS1 $r_{P1}$ and $i_{P1}$ bands between 0.3 and 1.0 mag.

@@

\begin{figure}
\plotone{astrometric_residuals.pdf}
\caption{
This plot illustrates the systematic offset in declination ($\Delta Decl$) of
objects observed in the POSS-II Blue epoch of the USNO-B catalog (plate 799), as
a function of the $g_{P1}$-band magnitude. For clarity, only a subset of objects
are plotted. The solid yellow circles show the median $\Delta Decl$ in magnitude
bins, and the solid red circles show the root-mean-square (rms) scatter in
magnitude bins. Note how the brighter objects are systematically offset by
$\sim200$ mas from the fainter objects. The rms scatter indicates that the
systematic precision in this coordinate and epoch is $\sim120$ mas.
\label{usnob_crossvalidation}}
\end{figure}

@@

Finally, to measure the proper motion of an object we fit a straight line to all
the available positions (max.~7). The proper motion of confirmed Ophiuchus
stream members is listed in Table~\ref{table1}.

@@

\subsection{Chemical abundances\label{abundances}}

The preliminary chemical abundances of five (RGB) stars in the Ophiuchus stream
(that is, the $v_{los}\sim290$ km s$^{-1}$ group), and observed by Hectochelle, 
are listed in Table~\ref{table_abundances}. The uncertainties of the determined
parameters are listed in the notes of Table~\ref{table_abundances}.

\begin{deluxetable}{lrrrrrr}
\tabletypesize{\scriptsize}
\setlength{\tabcolsep}{0.02in}
\tablecolumns{7}
\tablewidth{0pc}
\tablecaption{Chemical abundances of Ophiuchus stream stars\label{table_abundances}}
\tablehead{
\colhead{Name} & \colhead{$T_{eff}$} & \colhead{$\log g$} &
\colhead{${\rm [Fe/H]}$} & \colhead{${\rm [Mg/Fe]}$} &
\colhead{${\rm [Ca/Fe]}$} & \colhead{${\rm [Ti/Fe]}$}  \\
\colhead{ } & \colhead{(K)} & \colhead{(dex)} &
\colhead{(dex)} & \colhead{(dex)} & \colhead{(dex)} &
\colhead{(dex)}
}
\startdata
rgb1 & 5680 & 3.0 & -1.95 & -0.2 & 0.6 & 0.6 \\
rgb2 & 5450 & 2.8 & -2.02 &  0.1 & 0.6 & 0.3 \\
rgb3 & 5700 & 3.0 & -1.95 &  0.2 & 0.3 & 0.5 \\
rgb4 & 5500 & 2.8 & -1.95 &  0.3 & 0.6 & 0.6 \\
rgb5 & 5720 & 3.4 & -1.90 &  0.0 & 0.7 & 0.5 
\enddata
\tablecomments{The uncertainty in $T_{eff}$ is $<200$ K, $<0.4$ dex for
$\log g$, $\lesssim0.2$ dex for ${\rm [Fe/H]}$, and $\sim0.3$ dex for abundances
of $\alpha-$elements.}
\end{deluxetable}

@@

Based on spectroscopic data, the Ophiuchus stream is metal-poor
(${\rm [Fe/H]=-1.95\pm0.05}$ dex) and $\alpha-$enhanced
(${\rm [\alpha/Fe]=0.4\pm0.1}$ dex). As shown by \citet{scs93}, the
$\alpha-$enhanced stellar population models are equivalent to scaled-solar ones 
with the same global metal content ${\rm [M/H]}$, where ${\rm [M/H]}$ for
$\alpha-$enhanced models can be calculated using their Equation 3
\begin{equation}
    [M/H]\approxeq [Fe/H] + \log_{10}(0.638\times10^{[\alpha/Fe]} + 0.362).
\end{equation}
For the chemical abundance of the Ophiuchus stream, the above equation implies
${\rm [M/H]}=-1.7\pm0.2$ dex. This means that we should adopt the normal
distribution $\mathcal{N}(\log_{10}(Z/Z_\sun)| -1.7, 0.2)$ as the prior
probability of metallicity $Z$ (where $Z_\sun=0.0152$ is the solar metal content
used by this perticular set of PARSEC isochrones). However, in the context of
cross-validating our analysis, we decided to replace the above metallicity prior
in favor of a (less informative) prior that is uniform in the
$0.0001 < Z < 0.0004$ range. Even though a less informative prior was adopted,
at the end of Section~\ref{CMD_posterior} we find a very impressive consistency
between the posterior distribution of metallicity $Z$ (obtained using CMD
fitting) and the spectroscopic estimate of $Z$ (see bottom panel of
Figure~\ref{DM_Z_posteriors}). In the end, it is important to note that our
results would not have changed significantly if we used the more informative
prior for metallicity content $Z$.

@@

Fortunately, we do not need to rely only on confirmed members and can instead
use a much larger sample of stars in the vicinity of the Ophiuchus stream to
constrain its proper motion and extent. As we detail below, we use a
probabilistic approach (see Sections~\ref{vlos} and~\ref{CMD}) and model the
distribution of stars simultaneously in coordinate, proper motion, and
color-magnitude space as a mixture of stream and field (i.e., non-stream) stars.
Even though we do not {\em a priori} know which star is a true member of the
stream, we assume that as an ensemble, the stream stars have certain
characteristics which make them distinguishable from field stars (e.g., common
proper motion, distance, position on the sky and in the CMD), and that the
{\em scatter} in these characteristics is sufficiently small to overcome the
fact that there are a lot more field than stream stars. The narrow width of
the stream in color-magnitude (Figure~\ref{CMD_models}) and coordinate space
(Figure~1 of \citealt{ber14b}) support this assumption. After all, if the stream
did not have these characteristics, it likely would not have been detected by
\citet{ber14b} in the first place.

@@

In galactic coordinates, the distribution of stream stars is modeled as a
parabola with a Gaussian width $\sigma_b$ in the latitude direction, that is
\begin{equation}
    \mathcal{L}(l_k, b_k | \theta^{spatial}_{stream}) = \mathcal{N}(b_k | \nu(l_k), \sigma_b),
\end{equation}
where $\nu(l_k | A, B, C)=A + B(l_k-5) + C(l_k-5)^2$ is the galactic latitude of
the equator of the stream. The spatial distribution of field stars is modeled as
a straight line with a Gaussian width $\sigma^\prime_b$
\begin{equation}
    \mathcal{L}(l_k, b_k | \theta^{spatial}_{field}) = \mathcal{N}(b_k | \nu^\prime(l_k), \sigma^\prime_b),
\end{equation}
where $\nu^\prime(l_k | A^\prime, B^\prime)=A^\prime + B^\prime(l_k-5)$.

@@

\begin{figure}
\plotone{gl_vs_gb_track.pdf}
\caption{
The extent of the Ophiuchus stream in galactic coordinates. The thick solid line
shows the most probable model ($b(l)=31.369-0.73(l-5)-0.08(l-5)^2$ deg) for the
equator of the stream. To illustrate the uncertainty in the most probable
model, the thin semi-transparent red lines show 200 models sampled from the
posterior distribution. The vertical dashed lines show the likely extent of the
stream (see Section~\ref{proper_motion_extent}). The yellow points show the
positions of confirmed members and the arrow indicates the direction of movement
of the stream. The pixels show the probability-weighted number density map of
region, smoothed using a $6\arcmin$-wide Gaussian filter.
\label{gl_vs_gb_track}}
\end{figure}

We find the stream to be confined between galactic longitudes of 4.1 deg and 5.8
deg (Figure~\ref{gl_vs_gb_track}). When combined with the distance of the
stream (Figure~\ref{gl_vs_distance}), this result implies that the
deprojected length of the stream is $1.6\pm0.3$ kpc. Thus, the stream is very
foreshortened in projection, by a ratio of $6:1$. The galactic latitude of the
equator of the stream is at $b_{stream}(l) = 31.369 - 0.73(l-5) - 0.08(l-5)^2$
deg, and the width of the stream is $\sigma_b=6\pm1$ arcmin (in the galactic
latitude direction). In direction perpendicular to the stream's equator, the
stream is $\sim3.5\arcmin$ wide, which is similar to width of $\sim3\arcmin$
measured by \citet{ber14b}. Since the stream is very foreshortened, the
conversion and deprojection of the angular width is not a trivial matter, and
for this reason, we choose not provide an estimate of the physical width of the 
stream.

@@

\begin{figure}
\plotone{gl_vs_pm.pdf}
\caption{
Proper motion of the Ophiuchus stream as a function of galactic longitude $l$,
in the longitude ({\em top}) and latitude directions ({\em bottom}). The thick
solid lines show the most probable models ($\mu_l(l)=-2.6(l-5)-6.0$ mas
yr$^{-1}$, $\mu_b(l)=2.3(l-5)+2.3$ mas yr$^{-1}$). To illustrate the uncertainty
in most probable models, the thin semi-transparent red lines show 200 models
sampled from respective posterior distributions. For comparison, the
semi-transparent gray lines show the proper motion of field stars. The vertical 
dashed lines show the likely extent of the stream (see
Section~\ref{proper_motion_extent}).
\label{gl_vs_pm}}
\end{figure}

@@

\begin{figure*}
\plottwo{gl_vs_rv_orbit.pdf}{gl_vs_distance_orbit.pdf}
\plottwo{gl_vs_gb_track_orbit.pdf}{gl_vs_pm_orbit.pdf}
\caption{
This plot compares line-of-sight velocities ({\em top left}), distances
({\em top right}), positions ({\em bottom left}), and proper motions
({\em bottom right}) calculated by {\em galpy} (thin blue lines) with models
derived from observations (thin red lines). The observed and calculated values
are consistent within uncertainties, with proper motions being the only
exception.
\label{galpy_orbit_comparison}}
\end{figure*}

@@

\section{Time of disruption}

@@

We find that the tidal force is the strongest during pericenter+disk passages,
and the progenitor of the stream could have been disrupted during one of those
passages. To find if the progenitor was disrupted during one of these passages,
we can use {\em galpy}.

@@

The number density map of the N-body stream created using gyrfalcON is shown in
the bottom panel of Figure~\ref{observed_vs_mock_comparison}. The width and the
length of the N-body stream match the observed stream fairly well. For
comparison, if the cluster is evolved starting from the apocenter at
$t\sim-870$ Myr, the resulting stream is much longer and inconsistent with
observations. Based on this more realistic simulation, we conclude that the
Ophiuchus stream likely formed about 250 Myr ago and that its progenitor was
fully disrupted during a single disk+pericenter passage. If this scenario is
correct, then the answer to the question ``Where is the progenitor of the
Ophiuchus stream?'' is simple -- the Ophiuchus stream is all that is left of
the progenitor.

>>>>

\author{Branimir Sesar\altaffilmark{1,2}}
\author{Jo Bovy\altaffilmark{3,4}}
\author{Edouard J.~Bernard\altaffilmark{5}}
\author{Nelson Caldwell\altaffilmark{6}}
\author{Judith G.~Cohen\altaffilmark{7}}
\author{Morgan Fouesneau\altaffilmark{1}}
\author{Christian I.~Johnson\altaffilmark{6}}
\author{Melissa Ness\altaffilmark{1}}
%\author{Maria Bergemann\altaffilmark{1}}
\author{Annette M.~N.~Ferguson\altaffilmark{5}}
\author{Nicolas F.~Martin\altaffilmark{8,1}}
\author{Hans-Walter Rix\altaffilmark{1}}
\author{Edward F.~Schlafly\altaffilmark{1}}
%\author{builders}
\author{William S.~Burgett\altaffilmark{15}}
\author{Kenneth C.~Chambers\altaffilmark{9}} 
%\author{L.~Denneau\altaffilmark{9}} 
%\author{P.~Draper\altaffilmark{10}} 
%\author{H.~Flewelling\altaffilmark{9}} 
%\author{T.~Grav\altaffilmark{11}} 
%\author{J.~N.~Heasley\altaffilmark{9}} 
\author{Klaus W.~Hodapp\altaffilmark{9}} 
%\author{M.~E.~Huber\altaffilmark{9}} 
%\author{R.~Jedicke\altaffilmark{9}} 
\author{Nick Kaiser\altaffilmark{9}} 
%\author{R.-P.~Kudritzki\altaffilmark{9}}
%\author{G.~A.~Luppino\altaffilmark{9}}
%\author{R.~H.~Lupton\altaffilmark{12}}
\author{Eugene A.~Magnier\altaffilmark{9}}
\author{Imants Platais\altaffilmark{11}}
%\author{N.~Metcalfe\altaffilmark{10}}
%\author{D.~G.~Monet\altaffilmark{13}}
%\author{J.~S.~Morgan\altaffilmark{9}}
%\author{P.~M.~Onaka\altaffilmark{9}}
%\author{P.~A.~Price\altaffilmark{12}}
%\author{C.~W.~Stubbs\altaffilmark{14}}
%\author{W.~Sweeney\altaffilmark{9}}
\author{John L.~Tonry\altaffilmark{9}}
%\author{R.~J.~Wainscoat\altaffilmark{9}}
\author{Christopher Waters\altaffilmark{9}}
\author{Rosemary~F.~G.~Wyse\altaffilmark{11}}

@@

\begin{abstract}
The Ophiuchus stream is the most recently discovered stellar tidal stream in the
Milky Way \citep{ber14b}. We present high-quality spectroscopic data for 14
stream member stars obtained using the Keck and MMT telescopes. We confirm the
stream as a fast moving ($v_{los}\sim290$ km s$^{-1}$), kinematically-cold group
($\sigma_{v_{los}}\lesssim1$ km s$^{-1}$) of $\alpha-$enhanced and metal-poor
stars (${\rm [\alpha/Fe]\sim0.4}$ dex, ${\rm [Fe/H]\sim-2.0}$ dex). Using a
probabilistic technique, we model the stream simultaneously in line-of-sight
velocity, color-magnitude, coordinate, and proper motion space, and so determine
its distribution in 6D phase-space. We find that that the stream extends in
distance from 8 to 9.5 kpc from the Sun; it is 50 times longer than wide, merely
appearing highly foreshortened in projection. The analysis of the stellar
population contained in the stream suggests that it is $\sim13$ Gyr old, and
that its initial stellar mass was $\sim2\times10^4$ $M_\sun$ (or at least
$\ga4\times10^3$ $M_\sun$). Assuming a fiducial Milky Way potential, we fit an
orbit to the stream which matches the observed phase-space distribution, except
for some tension in the proper motions: the stream has an orbital period of
$\sim360$ Myr, and is on a fairly eccentric orbit ($e\sim0.68$) with a
pericenter of $\sim3.5$ kpc and an apocenter of $\sim17.5$ kpc. The phase-space
structure and stellar population of the stream show that its progenitor must
have been a globular cluster that was disrupted only $\sim250$ Myr ago. We do
not detect any significant overdensity of stars along the stream that would
indicate the presence of a progenitor, and conclude that the stream is all that
is left of the progenitor.
\end{abstract}

@@

To address the above questions, we need to know the orbit of the Ophiuchus
stream, and to determine its orbit we need to measure the stream's line-of-sight
velocity, distance, and proper motion. In Section~\ref{data}, we describe the
data we use in this work; the PS1 photometry and astrometry, the spectroscopic
follow-up of candidate stream members, and the measurement of their
line-of-sight velocities, element abundances, and proper motions. In
Section~\ref{results}, we provide a detailed characterization of the stream in
position, velocity, and abundance (7-D) phase space. The constraints obtained in
Section~\ref{results} are then used to constrain and examine the orbit of the
stream (Section~\ref{orbit}) and the time of disruption of its progenitor
(Section~\ref{disruption}). In Section~\ref{conclusions}, we discuss the nature
of the stream's peculiar orbit, highlight the solved and uncovered puzzles
related to the stream, and finally present our conclusions.

@@

\subsection{Element abundances}

Even though the primary goal of spectroscopic observations was to obtain precise
radial velocities, the wavelength range and the resolution of Hectochelle
spectra are sufficient to allow estimates of element abundances.

We determined stellar parameters from the continuum-normalised, radial
velocity-corrected spectra using the SMH code of \citet{cas14}, which is built
on the MOOG code of \citet{sne73}. Kurucz $\alpha-$enhanced (0.4 dex) model
atmospheres \citep{ck04} and a line list compiled from \citet{fre10} by
\citet{cas14} were used (see Table~\ref{lines}). First, effective temperatures
were calculated from 2MASS photometry and color-temperature calibrations of
\citet{ghb09} and spectroscopic temperatures were optimized around this value
using the SMH code, by removing abundance trends with line excitation potential.
Parameters of $\log g$ and ${\rm [Fe/H]}$ were determined using 14 Fe I and 2-3 
Fe II lines to achieve ionization balance, and microturbulence was calculated by
removing abundance trends with reduced equivalent width of the lines. Estimates 
of the $\alpha-$enhancement were obtained using only the few available clean Mg,
Ca and Ti (I and II) lines, which comprised a total of 6-8 lines per star. The
abundances were measured from equivalent widths and the lines we used are listed
in Table~\ref{lines}. Because Mg lines are strong and may be saturated, the
values of ${\rm [Mg/Fe]}$ are significantly different than values of
${\rm [Ca/Fe]}$ and ${\rm [Ti/Fe]}$.

@@

Proper motions are crucial constraints for determining the orbit of a stream
(e.g., \citealt{kop10}) To measure the proper motion of stars in the vicinity of
the Ophiuchus stream, we combine the astrometry provided by USNO-B \citep{mon03}
and 2MASS \citep{skr06} catalogs with the PS1 catalog. The USNO-B catalog lists
photometry and astrometry measured from photographic plates in five different
band-passes ($O$, $E$, $J$, $F$, and $N$). The USNO-B plates were exposed at
different epochs, and thus each object in the catalog can have a maximum of five
recorded positions. The 2MASS catalog provides only one position entry per
object.

To reduce the systematic offsets in astrometry between different catalogs, we
first calibrate USNO-B and 2MASS positions to a reference frame that is defined
by positions of galaxies observed in PS1 (which are on the ICRS coordinate
system). We define galaxies as objects that have the difference between
Point-Spread-Function (PSF) and aperture magnitudes in PS1 $r_{P1}$ and $i_{P1}$
bands between 0.3 and 1.0 mag.

@@

\begin{figure}
\plotone{astrometric_residuals.pdf}
\caption{
This plot illustrates the systematic offset in declination ($\Delta Decl$) of
objects observed in the POSS-II Blue epoch of the USNO-B catalog (plate 799), as
a function of the $g_{P1}$-band magnitude. For clarity, only a subset of objects
are plotted. The solid yellow circles show the median $\Delta Decl$ in magnitude
bins, and the solid red circles show the root-mean-square (rms) scatter in
magnitude bins. Note how the brighter objects are systematically offset by
$\sim200$ mas from the fainter objects. The rms scatter indicates that the
average precision in this coordinate and epoch is $\sim120$ mas for objects
brighter than $g_{P1}=19$.
\label{usnob_crossvalidation}}
\end{figure}

@@

Finally, to measure the absolute proper motion of an object we fit a straight
line to all available positions (min.~3, max.~7) as a function of time (the
epoch baseline is 58 years). The proper motion of confirmed Ophiuchus stream
members is listed in Table~\ref{table1}.

@@

\subsection{Element abundances\label{abundances}}

The preliminary element abundances of five (RGB) stars in the Ophiuchus stream
(that is, the $v_{los}\sim290$ km s$^{-1}$ group), and observed by Hectochelle, 
are listed in Table~\ref{table_abundances}. The uncertainties of the determined
parameters are listed in the notes of Table~\ref{table_abundances}.

\begin{deluxetable}{lrrrrrr}
\tabletypesize{\scriptsize}
\setlength{\tabcolsep}{0.02in}
\tablecolumns{7}
\tablewidth{0pc}
\tablecaption{Element abundances of Ophiuchus stream stars\label{table_abundances}}
\tablehead{
\colhead{Name} & \colhead{$T_{eff}$} & \colhead{$\log g$} &
\colhead{${\rm [Fe/H]}$} & \colhead{${\rm [Mg/Fe]}$} &
\colhead{${\rm [Ca/Fe]}$} & \colhead{${\rm [Ti/Fe]}$}  \\
\colhead{ } & \colhead{(K)} & \colhead{(dex)} &
\colhead{(dex)} & \colhead{(dex)} & \colhead{(dex)} &
\colhead{(dex)}
}
\startdata
rgb1 & 5680 & 3.0 & -1.95 & -0.2 & 0.6 & 0.6 \\
rgb2 & 5450 & 2.8 & -2.02 &  0.1 & 0.6 & 0.3 \\
rgb3 & 5700 & 3.0 & -1.95 &  0.2 & 0.3 & 0.5 \\
rgb4 & 5500 & 2.8 & -1.95 &  0.3 & 0.6 & 0.6 \\
rgb5 & 5720 & 3.4 & -1.90 &  0.0 & 0.7 & 0.5 
\enddata
\tablecomments{The uncertainty in $T_{eff}$ is $<200$ K, $<0.4$ dex for
$\log g$, $\lesssim0.2$ dex for ${\rm [Fe/H]}$, and $\sim0.3$ dex for abundances
of $\alpha-$elements.}
\end{deluxetable}

@@

Based on spectroscopic data, the Ophiuchus stream is metal-poor
(${\rm [Fe/H]=-1.95\pm0.05}$ dex) and $\alpha-$enhanced
(${\rm [\alpha/Fe]=0.4\pm0.1}$ dex). As shown by \citet{scs93}, the
$\alpha-$enhanced stellar population models are equivalent to scaled-solar ones 
with the same global metal content ${\rm [M/H]}$, where ${\rm [M/H]}$ for
$\alpha-$enhanced models can be calculated using their Equation 3
\begin{equation}
    [M/H]\approxeq [Fe/H] + \log_{10}(0.638\times10^{[\alpha/Fe]} + 0.362).
\end{equation}
For the element abundance of the Ophiuchus stream, the above equation implies
${\rm [M/H]}=-1.7\pm0.2$ dex. This means that we should adopt the normal
distribution $\mathcal{N}(\log_{10}(Z/Z_\sun)| -1.7, 0.2)$ as the prior
probability of metallicity $Z$ (where $Z_\sun=0.0152$ is the solar metal content
used by this perticular set of PARSEC isochrones). However, in the context of
cross-validating our analysis, we decided to replace the above metallicity prior
in favor of a (less informative) prior that is uniform in the
$0.0001 < Z < 0.0004$ range. Even though a less informative prior was adopted,
at the end of Section~\ref{CMD_posterior} we find a very impressive consistency
between the posterior distribution of metallicity $Z$ (obtained using CMD
fitting) and the spectroscopic estimate of $Z$ (see bottom panel of
Figure~\ref{DM_Z_posteriors}). In the end, it is important to note that our
results would not have changed significantly if we used the more informative
prior for metallicity content $Z$.

@@

Fortunately, we do not need to rely only on confirmed members and can use a much
larger sample of stars in the vicinity of the Ophiuchus stream to constrain its
proper motion and extent. As we detail below, we use a probabilistic approach
(see Sections~\ref{vlos} and~\ref{CMD}) and model the distribution of stars
simultaneously in coordinate, proper motion, and color-magnitude space as a
mixture of stream and field (i.e., non-stream) stars. Even though we do not
{\em a priori} know which star is a true member of the stream, we assume that as
an ensemble, the stream stars have certain characteristics which make them
distinguishable from field stars (e.g., common proper motion, distance, position
on the sky and in the CMD), and that the {\em scatter} in these characteristics
is sufficiently small to overcome the fact that there are a lot more field than
stream stars. The narrow width of the stream in color-magnitude
(Figure~\ref{CMD_models}) and coordinate space (Figure~1 of \citealt{ber14b})
support this assumption. After all, if the stream did not have these
characteristics, it likely would not have been detected by \citet{ber14b} in the
first place.

Even though the probabilistic approach we describe below uses all of the stars
in the vicinity of the Ophiuchus stream to constrain the its extent and proper
motion, we expect that most of the signal will come from main sequence turn-off
(MSTO) stars associated with the stream. As shown in Figure~\ref{CMD_models},
the stream's MSTO is bluer than the field population, which means that stars in
this region of the color-magnitude diagram are much more likely to be associated
with the stream than with the field population. Thus, the statistical weight of
such stars will be greater than, for example, the weight of stream's RGB stars,
which occupy the region of the color-magnitude diagram that is heavily dominated
by field stars.

@@

In galactic coordinates, the distribution of stream stars is modeled as a
parabola along the stream, with a Gaussian width $\sigma_b$ in the latitude
direction, that is
\begin{equation}
    \mathcal{L}(l_k, b_k | \theta^{spatial}_{stream}) = \mathcal{N}(b_k | \nu(l_k), \sigma_b),
\end{equation}
where $\nu(l_k | A, B, C)=A + B(l_k-5) + C(l_k-5)^2$ is the galactic latitude of
the equator of the stream. The spatial distribution of field stars is modeled as
a straight line with a Gaussian width $\sigma^\prime_b$
\begin{equation}
    \mathcal{L}(l_k, b_k | \theta^{spatial}_{field}) = \mathcal{N}(b_k | \nu^\prime(l_k), \sigma^\prime_b),
\end{equation}
where $\nu^\prime(l_k | A^\prime, B^\prime)=A^\prime + B^\prime(l_k-5)$.

@@

\begin{figure}
\plotone{gl_vs_gb_track.pdf}
\caption{
The extent of the Ophiuchus stream in galactic coordinates. The thick solid line
shows the most probable model ($b(l)=31.369-0.73(l-5)-0.08(l-5)^2$ deg) for the
equator of the stream. To illustrate the uncertainty in the most probable
model, the thin semi-transparent red lines show 200 models sampled from the
posterior distribution. The vertical dashed lines show the likely extent of the
stream (see Section~\ref{proper_motion_extent}). The yellow points show the
positions of confirmed members and the arrow indicates the direction of movement
of the stream. The gray-scale map shows the probability-weighted number density
of likely Ophiuchus stream members, smoothed using a $6\arcmin$-wide Gaussian
filter.
\label{gl_vs_gb_track}}
\end{figure}

We find the stream to be confined between galactic longitudes of 4.1 deg and 5.8
deg (Figure~\ref{gl_vs_gb_track}). When combined with the distance of the
stream (Figure~\ref{gl_vs_distance}), this result implies that the
deprojected length of the stream is $1.6\pm0.3$ kpc. Thus, the stream is very
foreshortened in projection, by a ratio of $6:1$. The galactic latitude of the
equator of the stream is at $b_{stream}(l) = 31.369 - 0.73(l-5) - 0.08(l-5)^2$
deg, and the width of the stream is $\sigma_b=6\pm1$ arcmin (in the galactic
latitude direction). In direction perpendicular to the stream's equator, the
stream is $\sim3.5\arcmin$ wide, which is similar to width of $\sim3\arcmin$
measured by \citet{ber14b}.

@@

\begin{figure}
\plotone{gl_vs_pm.pdf}
\caption{
Proper motion of the Ophiuchus stream, inferred from the ensemble of likely
stream members. The panels show the proper motion in galactic longitude
({\em top}) and latitude directions ({\em bottom}), as a function of galactic
longitude $l$. The thick solid lines show the most probable models
($\mu_l(l)=-2.6(l-5)-6.0$ mas yr$^{-1}$, $\mu_b(l)=2.3(l-5)+2.3$ mas yr$^{-1}$).
To illustrate the uncertainty in most probable models, the thin semi-transparent
red lines show 200 models sampled from respective posterior distributions. For
comparison, the semi-transparent gray lines show the proper motion of field
stars. The vertical dashed lines show the likely extent of the stream (see
Section~\ref{proper_motion_extent}).
\label{gl_vs_pm}}
\end{figure}

@@

\begin{figure*}
\plottwo{gl_vs_rv_orbit.pdf}{gl_vs_distance_orbit.pdf}
\plottwo{gl_vs_gb_track_orbit.pdf}{gl_vs_pm_orbit.pdf}
\caption{
This plot compares line-of-sight velocities ({\em top left}), distances
({\em top right}), positions ({\em bottom left}), and proper motions
({\em bottom right}) calculated by {\em galpy} (thin blue lines) with models
derived from observations (thin red lines). The observed and calculated values
are consistent within uncertainties, with the exception of the proper motions,
where the model cannot match the apparent gradient of the proper motions along
the stream.
\label{galpy_orbit_comparison}}
\end{figure*}

@@

\section{Time of disruption}\label{disruption}

@@

We find that the tidal force is the strongest during pericenter+disk passages,
and that the progenitor of the stream could have been disrupted during one of
those passages. To find if the progenitor was disrupted during one of these
passages, we can use {\em galpy}.

@@

The number density map of the N-body stream created using gyrfalcON is shown in
the bottom panel of Figure~\ref{observed_vs_mock_comparison}. The width and the
length of the N-body stream match the observed stream fairly well. For
comparison, if the cluster is evolved starting from the apocenter at
$t\sim-870$ Myr, the resulting stream is much longer and inconsistent with
observations. Based on this more realistic simulation, we conclude that the
Ophiuchus stream likely formed about 250 Myr ago and that its progenitor was
fully disrupted during a single disk+pericenter passage.
